---
title: "GA4GH/DREAM Submission Summary"
output: 
  html_notebook:
    code_folding: "hide"
---

```{r, message=FALSE, echo=FALSE}
library(synapseClient)
library(tidyverse)
library(stringr)
library(ggplot2)
library(forcats)
library(viridis)
library(plotly)
```

Submission data — aggregated in the Synapse table `syn103088233` — was downloaded using the `synapseClient` package for R on `r lubridate::now("UTC")` UTC. Note that this table includes only submissions for which both `status` and `reportStatus` are **"VALIDATED"**.

```{r, message=FALSE, include=FALSE}
synapseLogin()
submission_table <- synTableQuery("select * from syn10388233")
valid_df <- submission_table@values
```

Basic cleaning of platform names is also handled here. Note that while the `platform` column in the source table is parsed directly from participant-provided documentation (i.e., the **README** for each submission), `platformHandle` represents the cleaned or merged platform names.

```{r}
valid_clean_df = valid_df %>% 
    rownames_to_column("row_id") %>% 
    mutate(platformHandle = case_when(
               str_to_lower(platform) == "dockstore cli" ~ "dockstore",
               platform == "Toil (new cwltoil branch)" ~ "Toil",
               str_to_lower(platform) == "toil" ~ "Toil",
               TRUE ~ as.character(platform)
           )
    ) %>% 
    column_to_rownames("row_id")
```

As of the latest update, there have been:

+ **`r nrow(valid_clean_df)`** total successful submissions...
+ by **`r n_distinct(valid_clean_df$team)`** different teams or individuals...
+ run on **`r n_distinct(valid_clean_df$platformHandle)`** different platforms.

```{r, include=FALSE}
if (!all_equal(valid_clean_df, submission_table@values) == TRUE) {
    schema <- synGet("syn10388233")
    if (!"platformHandle" %in% map_chr(schema@columns@content, "name")) {
        new_col <- TableColumn(name = "platformHandle", columnType = "STRING")
        schema <- synAddColumn(schema, new_col)
        schema <- synStore(schema)
    }
    submission_table@values <- valid_clean_df
    submission_table <- synStore(submission_table)
}
```

# Submission breakdown

Plots below were generated using `ggplot2` and `plotly`. Colors indicate the workflow that was run for each submission.

### Successful submissions per platform

Participants should nominally run each workflow using their platform of choice. We then extract platform information that the participant has recorded (among other YAML-style fields) in the report / **README** wiki for each submission. Each row in the plot below shows the number of validated and documented submissions for a particular workflow execution platform.

```{r subs_by_platform}
subs_by_platform <- valid_clean_df %>% 
    mutate(platformHandle = fct_rev(platformHandle)) %>% 
    ggplot(aes(platformHandle)) +
    geom_bar(aes(fill = workflow), alpha = 0.8, colour = "white",
             position = position_stack(reverse = TRUE)) +
    scale_fill_viridis(discrete = TRUE) + 
    coord_flip() +
    xlab("") +
    ylab("") +
    theme(plot.title = element_text(face = "bold"),
          legend.title = element_blank())

ggplotly(subs_by_platform, width = 800) %>%
    layout(margin = list(l = 100, r = 250))
```


### Successful submissions per team/participant

Participants can submit as an individual or as part of a team. We record this information when validating each submission. Each row in the plot below shows the number of validated and documented submissions for a team or individual.

```{r subs_by_team}
subs_by_team <- valid_clean_df %>% 
    mutate(workflow = ifelse(workflow == "encode_mapping",
                             "encode_mapping_workflow", workflow),
           team = fct_rev(team)) %>% 
    ggplot(aes(team)) +
    geom_bar(aes(fill = workflow), alpha = 0.8, colour = "white",
             position = position_stack(reverse = TRUE)) +
    scale_fill_viridis(discrete = TRUE) + 
    coord_flip() +
    xlab("") +
    ylab("") +
    theme(plot.title = element_text(face = "bold"),
          legend.title = element_blank())

ggplotly(subs_by_team, width = 800) %>%
    layout(margin = list(l = 280, r = 250))
```



---

# Session info

```{r}
sessionInfo()
```
