---
title: "GA4GH/DREAM Workflow Execution Challenge Submissions"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
---

```{r setup, include=FALSE}
library(flexdashboard)
library(synapseClient)
library(tidyverse)
library(stringr)
library(ggplot2)
library(forcats)
library(viridis)
library(plotly)

custom_theme_bw <- function() {
    theme_bw() +
        theme(axis.title = element_text(face = "bold"),
              legend.title = element_text(face = "bold"),
              plot.title = element_text(face = "bold"))
}

```

```{r, message=FALSE, include=FALSE}
synapseLogin()
submission_table <- synTableQuery("select * from syn11713824")
valid_clean_df <- submission_table@values
```


Row {data-height=150}
-----------------------------------------------------------------------

### **Overview**

Submission data — cleaned and aggregated in the Synapse table `syn11713824` — was downloaded using the `synapseClient` package for R on `r lubridate::now("UTC")` UTC. The code used for cleaning participant-provided information can be found [here](https://github.com/GA4GH-DREAM/ga4gh-dream-challenge/blob/master/submission_cleaning.Rmd)  (note: undedited submission information can be found in Synapse table `syn11700144`, and full participant documentation can be found through the [**Submission Tracker boards**](https://www.synapse.org/#!Synapse:syn8507133/wiki/451414) leaderboards).

-----

Note that successful submissions represent those submissions for which both `status` and `reportStatus` are **“VALIDATED”**.


Row
-----------------------------------------------------------------------

### Workflows ran

```{r}
workflows <- n_distinct(valid_clean_df$workflow)
valueBox(workflows, icon = "fa-cogs")
```

### Platforms represented

```{r}
platforms <- n_distinct(valid_clean_df$platform)
valueBox(platforms, icon = "fa-wrench")
```



Row
-----------------------------------------------------------------------

### Fully Documented Submissions

```{r}
eval_ids <- unique(valid_clean_df$scopeId)
total_validated <- map(eval_ids, function(eval_id) {
    subs <- synGetSubmissions(eval_id, status = "VALIDATED", limit = 100)
    subs@totalNumberOfResults
}) %>% 
    reduce(sum)
total_documented <- nrow(valid_clean_df)
gauge(total_documented, min = 0, max = total_validated)
```

> (out of all submissions with validated outputs)

### Teams Successfully Submitted

```{r}
teams <- synRestGET('/challenge/3090/challengeTeam')
total_teams <- teams$totalNumberOfResults

successful_teams <- n_distinct(valid_clean_df$team)
gauge(successful_teams, min = 0, max = total_teams)
```

> (out of all registered Challenge teams)

### Individuals Successfully Submitted

```{r}
individuals <- synRestGET('/challenge/3090/participant')
total_individuals <- individuals$totalNumberOfResults

successful_individuals <- n_distinct(valid_clean_df$user)
gauge(successful_individuals, min = 0, max = total_individuals)
```

> (out of all registered Challenge participants)

Row  
-----------------------------------------------------------------------

### **Workflow:Platform Matrix**

```{r}
plot_workflow_matrix <- function(groupvar) {
    submission_df <- valid_clean_df
    y_vals <- unique(submission_df[["workflow"]])
    x_vals <- unique(submission_df[[names(groupvar)]])
    num_x <- length(x_vals)
    
    y_margin <- max(purrr::map_int(y_vals, stringr::str_length))
    x_margin <- max(purrr::map_int(x_vals, stringr::str_length))
    
    groupcol = as.name(names(groupvar))
    plot_df <- submission_df %>% 
        group_by(.dots = c("workflow", names(groupvar))) %>% 
        tally() %>% 
        mutate(label = glue::glue("<b><i>{workflow}</i></b><br>[{groupvar}] <b>{group}</b>: {count} submissions",
                                  workflow = workflow, 
                                  groupvar = as.character(groupvar),
                                  group = rlang::UQ(groupcol),
                                  count = n)) %>% 
        ungroup() %>% 
        mutate(workflow = forcats::fct_rev(workflow))

    
    p <- plot_df %>% 
        ggplot(aes_string(x = names(groupvar), y = "workflow", text = "label")) +
        geom_tile(aes(fill = n), alpha = 0.8, colour = "black", size = 0.2) +
        scale_fill_viridis(discrete = FALSE) +
        xlab("") +
        ylab("") +
        custom_theme_bw() +
        theme(plot.title = element_text(face = "bold"),
              legend.title = element_blank(),
              axis.text.x = element_text(angle = 45, hjust = 1),
              panel.grid.major = element_blank()
        )

    ggplotly(p, tooltip = "text") %>%
        layout(margin = list(l = 10 + y_margin * 6,
                             b = 10 + x_margin * 4.5),
               font = list(family = "Roboto, Open Sans, sans-serif"))
}
```

```{r, message=FALSE, warning=FALSE}
plot_workflow_matrix(groupvar = list(platform = "Platform"))
```

Row {data-height=50}
-----------------------------------------------------------------------

### **Submission Breakdowns**

Switch between tabs below to view the breakdown of successful submissions by platform, environment, or team.

Row  {.tabset .tabset-fade}
-----------------------------------------------------------------------

```{r, include=FALSE}
plot_workflow_breakdown <- function(groupvar) {
    submission_df <- valid_clean_df
    fill_vals <- unique(submission_df[["workflow"]])
    bar_vals <- unique(submission_df[[names(groupvar)]])
    num_bars <- length(bar_vals)
    
    fill_margin <- max(purrr::map_int(fill_vals, stringr::str_length))
    bar_margin <- max(purrr::map_int(bar_vals, stringr::str_length))
    
    groupcol = as.name(names(groupvar))
    plot_df <- submission_df %>% 
        group_by(.dots = c("workflow", names(groupvar))) %>% 
        tally() %>% 
        mutate(label = glue::glue("<b><i>{workflow}</i></b><br>[{groupvar}] <b>{group}</b>: {count} submissions",
                                  workflow = workflow, 
                                  groupvar = as.character(groupvar),
                                  group = rlang::UQ(groupcol),
                                  count = n))

    
    p <- plot_df %>% 
        ggplot(aes_string(x = names(groupvar), y = "n", text = "label")) +
        geom_col(aes(fill = workflow), alpha = 0.8, colour = "black", size = 0.2) +
        scale_fill_viridis(discrete = TRUE) +
        coord_flip() +
        xlab("") +
        ylab("") +
        custom_theme_bw() +
        theme(plot.title = element_text(face = "bold"),
              legend.title = element_blank()) +
        ggtitle(glue::glue("Workflows by {groupvar}", 
                           groupvar = as.character(groupvar)))

    ggplotly(p, tooltip = "text") %>%
        layout(margin = list(l = 12 + bar_margin * 6,
                             r = 10 + fill_margin * 6),
               font = list(family = "Roboto, Open Sans, sans-serif"))
}
```

### Platform

```{r, warning=FALSE, message=FALSE}
plot_workflow_breakdown(groupvar = list(platform = "Platform"))
```

### Environment

```{r, message=FALSE, warning=FALSE}
plot_workflow_breakdown(groupvar = list(environment = "Environment"))
```

### Team

```{r, message=FALSE, warning=FALSE}
plot_workflow_breakdown(groupvar = list(team = "Team"))
```


### Type

```{r, message=FALSE, warning=FALSE}
plot_workflow_breakdown(groupvar = list(workflow_type = "Type"))
```